---
title: "Reproducible Data Course Project #1"
author: "HCANG"
date: "April 18, 2019"
output: html_document
---

##Step 1:
####Code for reading and downloading the data set, and pre-loading necessary packages

```{r,echo=TRUE,message=FALSE}
library(dplyr)
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
file <- "repdata_2Fdata_Factivity.zip"
if (!file.exists(file)){
  download.file(fileURL,file,mode="wb")
  }
path <- "repdata_Factivity"
if(!file.exists(path)){
  unzip(file)
}
activityDF <- read.csv("activity.csv")
```

##Step 2:
####What is the mean total number of steps taken per day
```{r,echo=TRUE}
StepPerDay <- activityDF %>% group_by(date) %>% summarize(sum(steps,na.rm=TRUE))
colnames(StepPerDay) <- c("Date","Steps")
hist(StepPerDay$Steps,xlab="Number of Steps",ylab="Frequency",main="Steps Per Day")
```

##Step 3:
####Calculate the mean and median of the total number of steps taken
```{r,echo=TRUE}
MeanStepsPerDay <- mean(StepPerDay$Steps[StepPerDay$Steps!=0],na.rm=TRUE)
MedianStepsPerDay <- median(StepPerDay$Steps[(StepPerDay$Steps!=0)],na.rm=TRUE)
MeanStepsPerDay
MedianStepsPerDay
```

##Step 4:
####Creating a time series plot for daily activity (in 5 minute intervals)
```{r,echo=TRUE}
ActivityIntervalAve <- activityDF %>% group_by(interval) %>% summarize(mean(steps,na.rm=TRUE))
colnames(ActivityIntervalAve) <- c("Interval","Steps")
plot(x=ActivityIntervalAve$Interval,y=ActivityIntervalAve$Steps,type="l",main="Average Daily Steps per Interval")
```

##Step 5:
####Calculate which interval returns the most average number of steps
```{r,echo=TRUE}
ActivityIntervalAve[which(ActivityIntervalAve$Steps==max(ActivityIntervalAve$Steps)),]
```

##Step 6:
####Calculating total number of NA's in the data set
```{r,echo=TRUE}
sum(is.na(activityDF))
```

##Step 7:
####Replacing NA's with the mean for the each interval
```{r,echo=TRUE}
activityDFNA <- activityDF
for (i in 1:nrow(activityDFNA)){
   ifelse(is.na(activityDFNA$steps[i])==TRUE,
          activityDFNA$steps[i] <- ActivityIntervalAve$Steps[ActivityIntervalAve$Interval==activityDFNA$interval[i]],
          activityDFNA$steps[i] <- activityDF$steps[i])
}
```

##Step 8:
####Create a histogram for the total number of steps taken per day
```{r,echo=TRUE}
StepPerDayNA <- activityDFNA %>% group_by(date) %>% summarize(sum(steps,na.rm=TRUE))
colnames(StepPerDayNA) <- c("Date","Steps")
hist(StepPerDayNA$Steps,xlab="Number of Steps",ylab="Frequency",main="Steps Per Day")
```

##Step 9:
####Calculating the mean and mediam of the new data set
```{r,echo=TRUE}
MeanStepsPerDayNA <- mean(StepPerDayNA$Steps[StepPerDayNA$Steps!=0],na.rm=TRUE)
MedianStepsPerDayNA <- median(StepPerDayNA$Steps[(StepPerDayNA$Steps!=0)],na.rm=TRUE)
MeanStepsPerDayNA
MedianStepsPerDayNA
```
#####The mean remains the same while the median increased slightly. Imputing the mean doesn't really skew the data for this set

##Step 10:
####Creating a new factor variable for weekdays and weekend
```{r,echo=TRUE}
activityDFNA$TypeofDay <- ifelse(weekdays(as.Date(activityDFNA$date))=="Saturday"|weekdays(as.Date(activityDFNA$date))=="Sunday","Weekend","Weekday")
```

##Step 11:
####Creating a time series plot for weekdays and weekends
```{r,echo=TRUE}
par(mfrow=c(1,2))
ActivityIntervalAveTypeofDay <- activityDFNA %>% group_by(interval,TypeofDay) %>% summarize(mean(steps,na.rm=TRUE))
colnames(ActivityIntervalAveTypeofDay) <- c("Interval","TypeofDay","Steps")
ActivityIntervalAveWeekend <- ActivityIntervalAveTypeofDay[which(ActivityIntervalAveTypeofDay$TypeofDay=="Weekend"),]
ActivityIntervalAveWeekday <- ActivityIntervalAveTypeofDay[which(ActivityIntervalAveTypeofDay$TypeofDay=="Weekday"),]

plot(x=ActivityIntervalAveWeekend$Interval,y=ActivityIntervalAveWeekend$Steps,type="l",main="Average Daily Steps (Weekend)",xlab="Interval",ylab="Steps")
plot(x=ActivityIntervalAveWeekday$Interval,y=ActivityIntervalAveWeekday$Steps,type="l",main="Average Daily Steps (Weekday)",xlab="Interval",ylab="Steps")
```

